### 2021-11-16 documents/floats/say.tex
added yukpa ka, generating all tables again

\begin{table}
\centering
\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102; p.c., Spike Gildea]{mattei1994diccionario}}
\label{tab:say}
\begin{tabular}[t]{@{}llllll@{}}
\mytoprule
Language & \multicolumn{5}{l}{Form} \\
\midrule
\kaxui   &   \obj{ka[s]} &  k &  a &  s &    \\
\hixka   &   \obj{ka[h]} &  k &  a &  h &    \\
\waiwai  &   \obj{ka[s]} &  k &  a &  s &    \\
\arara   &      \obj{ke} &  k &  e &    &    \\
\ikpeng  &      \obj{ke} &  k &  e &    &    \\
\bakairi &      \obj{ke} &  k &  e &    &    \\
\trio    &      \obj{ka} &  k &  a &    &    \\
\akuriyo &      \obj{ka} &  k &  a &    &    \\
\carijo  &      \obj{ka} &  k &  a &    &    \\
\wayana  &   \obj{ka[i]} &  k &  a &    &  i \\
\apalai  &  \obj{ka[ʃi]} &  k &  a &  ʃ &  i \\
\kalina  &      \obj{ka} &  k &  a &    &    \\
\kapon   &      \obj{ka} &  k &  a &    &    \\
\pemon   &      \obj{ka} &  k &  a &    &    \\
\macushi &      \obj{ka} &  k &  a &    &    \\
\panare  &   \obj{ka[h]} &  k &  a &  h &    \\
\uxc     &      \obj{ki} &  k &  i &    &    \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{table}
\centering
\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
\label{tab:say}
\begin{tabular}[t]{@{}llllll@{}}
\mytoprule
Language & \multicolumn{5}{l}{Form} \\
\midrule
\kaxui   &   \obj{ka[s]} &  k &  a &  s &    \\
\hixka   &   \obj{ka[h]} &  k &  a &  h &    \\
\waiwai  &   \obj{ka[s]} &  k &  a &  s &    \\
\arara   &      \obj{ke} &  k &  e &    &    \\
\ikpeng  &      \obj{ke} &  k &  e &    &    \\
\bakairi &      \obj{ke} &  k &  e &    &    \\
\trio    &      \obj{ka} &  k &  a &    &    \\
\akuriyo &      \obj{ka} &  k &  a &    &    \\
\carijo  &      \obj{ka} &  k &  a &    &    \\
\wayana  &   \obj{ka[i]} &  k &  a &    &  i \\
\apalai  &  \obj{ka[ʃi]} &  k &  a &  ʃ &  i \\
\kalina  &      \obj{ka} &  k &  a &    &    \\
\kapon   &      \obj{ka} &  k &  a &    &    \\
\pemon   &      \obj{ka} &  k &  a &    &    \\
\macushi &      \obj{ka} &  k &  a &    &    \\
\panare  &   \obj{ka[h]} &  k &  a &  h &    \\
\uxc     &      \obj{ki} &  k &  i &    &    \\
\yukpa   &      \obj{ka} &  k &  a &    &    \\
\bottomrule
\end{tabular}
\end{table}

### 2021-11-16 documents/sections/verbs.tex
added yukpa ka, generating all tables again

\ex<kati> \apalai\\
\obj{kaʃi-ko} \qu{say!}\\
\parencite[35]{koehn1986apalai}
\xe
%
\cref{tab:say} shows a comparison of the longest attested forms for each language.
Cognate segments in \cref{tab:say,tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.
%
\input{floats/say.tex}%
%
This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}).
We do not know the first person form of its \yukpa reflex \obj{ka}.

@@

\pex<hixsay>\hixka
\a<hix-119>
\begingl
\glpreamble onɨ wyaro nkekonɨ bɨryekomo, tɨyonɨ wya//
\gla onɨ wjaro n-ka-jakonɨ bɨrʲekomo tɨ-jonɨ wja//
\glb this like \gl{3}-say-\gl{rem}.\gl{cont} boy \gl{cor}-mother \gl{obl}//
\glft \qu{This is what the boy said to his mother.} \parencite[][36]{hixkaryanaderby1985}//
\endgl
\a<ordpl-35>
\begingl
\glpreamble moro ha, ketxkoná hatá.//
\gla moro ha ka-jat͡ʃkonɨ hatɨ//
\glb \gl{med}.\gl{dem}.\gl{inan} \gl{ints} say-\gl{rem}.\gl{cont}.\gl{pl} \gl{hsy}//
\glft \qu{“That one there” they said.} \parencite[][14]{derbyshire1965textos}//
\endgl
\xe
In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}).
In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
At least the \trio cognate shows the same pattern, albeit inconsistently \parencite[267]{triocarlin2004}.

Comparative evidence also points to \rc{ka[ti]} \qu{to say} showing transitive traits:
\trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
The exceptionality of \obj{ka} \qu{to say} taking \obj{-po} \qu{\gl{caus}.\gl{tr}} has also been noted for \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
The agent nominalizer \rc{-ne} gave rise to the \panare gnomic verbal suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
The occurrence of \obj{-ɲe} on \obj{ka} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, contrastive with the intransitive analysis by \textcite[102]{mattei1994diccionario}.
Finally, reflexes of the causativizer \rc{-metipo}, usually restricted to transitive verbs \parencite{gildea2015valency}, are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.

>>>>

\ex<kati> \apalai\\
\obj{kaʃi-ko} \qu{say!}\\
\parencite[35]{koehn1986apalai}
\xe
%
\cref{tab:say} shows a comparison of the longest attested forms for each language.%
\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.}
%
\input{floats/say.tex}%
%
This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}).
We do not know the first person form of its \yukpa reflex \obj{ka}.

@@

\pex<hixsay>\hixka
\a<hix-119>
\begingl
\glpreamble onɨ wyaro nkekonɨ bɨryekomo, tɨyonɨ wya//
\gla onɨ wjaro n-ka-jakonɨ bɨrʲekomo tɨ-jonɨ wja//
\glb this like \gl{3}-say-\gl{rem}.\gl{cont} boy \gl{cor}-mother \gl{obl}//
\glft \qu{This is what the boy said to his mother.} \parencite[][36]{hixkaryanaderby1985}//
\endgl
\a<ordpl-35>
\begingl
\glpreamble moro ha, ketxkoná hatá.//
\gla moro ha ka-jat͡ʃkonɨ hatɨ//
\glb \gl{med}.\gl{dem}.\gl{inan} \gl{ints} say-\gl{rem}.\gl{cont}.\gl{pl} \gl{hsy}//
\glft \qu{“That one there” they said.} \parencite[][14]{derbyshire1965textos}//
\endgl
\xe
In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}).
In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
At least the \trio cognate shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.

Derivational suffixes also point to \rc{ka[ti]} \qu{to say} being transitive:
\trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
The exceptionality of \obj{ka} \qu{to say} taking \obj{-po} \qu{\gl{caus}.\gl{tr}} has also been noted for \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
The agent nominalizer \rc{-ne} gave rise to the \panare gnomic verbal suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
The occurrence of \obj{-ɲe} on \obj{ka} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, contrasting with the intransitive analysis by \textcite[102]{mattei1994diccionario}.
Finally, reflexes of the causativizer \rc{-metipo}, usually restricted to transitive verbs \parencite{gildea2015valency}, are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
